---
title: "Machinery COsts"
output: html_document
date: "2024-10-28"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set wd}
setwd("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/4R Advocate Case Studies/")
```

```{r packages}
library(here)
library(readxl)
library(tidyverse)
```


```{r DE 2024 Advocate}
advocate1 = read_excel("2024 Bryant Lowe/DE Inputs.xlsx")
```

```{r GA 2024 Advocate}
advocate1 = read_excel("2024 Mason Roberts/GA Inputs.xlsx", sheet = 2)
```

```{r load calculator sheets}
#Machinery 
tractors = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/Machinery.xlsx", sheet = 1)

machines = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/Machinery.xlsx", sheet = 2)

implements = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/Machinery.xlsx", sheet = 3)

equipment = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/Machinery.xlsx", sheet = 5)

irrigation = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/Machinery.xlsx", sheet = 4)

#Prices for Diesel
prices = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/FertilizerPrices_Dec2024.xlsx", sheet = 2)

```

# Custom Costs 

```{r Prepare Data Frame for Machinery Costs}
# Clean out empty columns 
df1 <- advocate1 %>%
  mutate(totalCustomCost = appCostPerAc) 

```

# Drop purchase price columns if using standard price for machinery 
```{r}
df2 <- df1 %>%
  # Check if 'machPriceType' is "standard" and drop the specified columns
  {if ("standard" %in% .$machPriceType) 
     select(., -machPurchPrice, -impPurchPrice) 
   else .}

```

# Machinery Self Application Costs 
Tact added to end of values associate with Tractors 
Imp added to end of columns associated with Implements 
Mach added to end of columns associated with Machines 

```{r}
# Join implement cost values 
df3 <- df2 %>%
  left_join(implements, by = c("implement", "impWdthFt"))
  # select(where(~ !all(is.na(.))))

# Join tractor costs values
df4 <- df3 %>%
  left_join(tractors, by = c("machine", "tractorHp"),
            suffix = c("Imp", "Trac"))  #Imp (for implement) added to original columns
  # %>% select(where(~ !all(is.na(.)))) #Drop columns that are all empty

# Join machine cost values
df5 <- df4 %>%
  left_join(machines, by = c("machine"), # Join tractor costs by its HP class
            suffix = c("", "Mach")) #Imp (for implement) added to original columns
```

# irrigation Costs 
```{r}
# Join irrigation cost values 
df6 <- df5 %>%
  left_join(irrigation, by = c("irrigation")) # Do not drop extra columns for calculation of cost as placeholder columns 
```

## Variable Costs 
### Labor Costs 
= labor($/hr) * Per Rate of implement (hr/ac)
1.	Per Rate (hr/ac) = 1/[speed * efficiency * (implement length/8.25)]

ii.	Center Pivot 
= labor($/hr) * labor (hrs/ac-in) * irrigation (ac-in) 

iii.	Drip irrigation 
= labor($/hr) * labor (hr/application) 

```{r Calculate Labor Costs}
# Cost for labor in $/hr 
Labor = 15 

df7 <- df6 %>%
  mutate(
    laborCost = case_when(
    is.na(irrigation) ~ perRateHrAc * Labor,
    irrigation == "centerPivot" ~ Labor * laborHrAcIn * irrVolAcIn, 
    irrigation == "dripIrrigation"  ~  Labor * laborHrsApp, 
    TRUE ~ 0  # Default case if none of the above conditions are met
  ))
```

### Fuel Costs 
i. Tractor + Implement or Self-propelled machine 
= fuel use(gal/hr) * fuel cost * Per Rate of implement (hr/ac) 

ii. Center Pivot  
= fuel($/gal) * fuel (gal/ac-in) * irrigation (ac-in)  

iii. Drip irrigation  
None 

```{r Fuel}
# fuel cost global variable for irrigation (not sure where it comes from - could use average?)**
fuel_cost2 = 2.75

# # Join prices df to get diesel costs for relevant years (3 yr avg)
# df8 <- df7 %>%
#   left_join(prices %>% select(Year, Diesel), by = c("year" = "Year"))

# Only join diesel prices
diesel <- prices %>%
  group_by(Year) %>%
  summarise(Diesel = mean(Diesel))
# Join prices df to get diesel costs for relevant years (3 yr avg)
df8 <- df7 %>%
  left_join(diesel, by = c("year" = "Year"))

# Calculate Fuel Costs 
df9 <- df8 %>%
  mutate(fuelCost = case_when(
    is.na(irrigation) ~ fuelUseGalHr * perRateHrAc * Diesel,  # If irrigation is NA
    irrigation == "centerPivot" ~ fuel_cost2 * fuelGalAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    TRUE ~ 0  # Default case if none of the above conditions are met
  ))


```

### R&M Cost  
i. Tractor + Implement or Self-propelled machine 
= (tractor purchase price * R&M / useful life /annual hours) * Per Rate implement (hr/ac) +  
(Implement purchase price * R&M / useful life / annual hours) * Per Rate implement (hr/ac) 

ii. Center Pivot  
= R&M($/ac-in) * irrigation (ac-in)
```{r R&M}
df10 <- df9 %>%
  mutate(rmCost = case_when(
    is.na(irrigation) ~ ((machPurchPrice * rmTrac/ usefulLifeYrsTrac / annualUseHrsTrac) * perRateHrAc + 
                           (impPurchPrice * rmImp /usefulLifeYrsImp / annualUseHrsImp) * perRateHrAc),
    irrigation == "centerPivot" ~ rmDolAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    TRUE ~ 0)  # Default case if none of the above conditions are met
  )
```

## Ownership Costs 

### Depreciation 
  
i. Machine/Implement  
= (((tractor price – (tractor price * salvage)) / useful life / annual hours) * per rate of implement) +  
((implement price – (implement price * salvage)) / useful life / annual hours * per rate of implement))  

ii. Center Pivot  
= Ownership cost * irrigation (ac-in)  

iii. Drip irrigation
= ownership cost * length (ft) 

```{r Depreciation}
df11 <- df10 %>%
  mutate(deprecCosts = case_when(
    is.na(irrigation) ~ ((((machPurchPrice - (machPurchPrice * salvageTrac))/
                           usefulLifeYrsTrac / annualUseHrsTrac) * perRateHrAc) +
                        (((impPurchPrice - (impPurchPrice * salvageImp)) / 
                           usefulLifeYrsImp / annualUseHrsImp) * perRateHrAc)),  # If irrigation is NA
    irrigation == "centerPivot" ~ 
      ownershipDolAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    irrigation == "dripIrrigation" & !is.na(irrigation) ~ 
      ((ownershipDolFt * impWdthFt) * (irrVolAcIn/totalWater)),
    TRUE ~ 0)) 
```
### Interest 

i. Machine/Implement 
= (((machinery price+(machinery price * salvage))/2* interest rate / annual hours * per rate of implement) + 
((implement price +(implement price * salvage))/2* interest rate / annual hours * per rate of implement))

ii. Center pivot  
None 

iii. Drip irrigation  
None

```{r Interest}
interest_rate = 0.05

df12 <- df11 %>%
  mutate(interestCosts = case_when(
    is.na(irrigation) ~ 
      (((machPurchPrice + (machPurchPrice * salvageTrac)) / 2 * 
          interest_rate / annualUseHrsTrac * perRateHrAc) +
         ((impPurchPrice+(impPurchPrice*salvageImp))/2*interest_rate/annualUseHrsImp *
                            perRateHrAc)),
    TRUE ~ 0) # Default case if none of the above conditions are met
) 

```

## Total Costs 
```{r Total Machinery Costs}
df13 <- df12 %>% 
  # mutate(# Add dummy columns for values in equations when farmer doesn't use their own machinery  
  #   if (!"TotalCustomCost" %in% colnames(df)) {df <- df %>% mutate(TotalCustomCost = NA)}) %>%
  mutate(variableCosts = 
           coalesce(laborCost, 0) + 
           coalesce(fuelCost,0) + 
           coalesce(rmCost, 0))%>%
  mutate(ownershipCosts = 
           coalesce(interestCosts, 0) + 
           coalesce(deprecCosts, 0)) %>%
  mutate(machineryCosts = 
           coalesce(variableCosts, 0) + 
           coalesce(ownershipCosts, 0) + 
           coalesce(totalCustomCost, 0))

  
inputs1 <- df13 %>%
  select(-c(29:68))

  # select(where(~ !all(is.na(.)))) 
  
```


# Write File 
```{r Write File}
# Save csv
write.csv(inputs1, file = "2024 Bryant Lowe/Outputs/DE Inputs_Costs.csv", row.names = FALSE)

write.csv(inputs1, file = "2024 Mason Roberts/Outputs/GA Inputs_Costs.csv", row.names = FALSE)
```

```{r}
# rm(list = ls())
```
