---
title: "Machinery COsts"
output:
  word_document: default
  html_document: default
date: "2024-10-28"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set wd}
setwd("C:/Users/LeannaNigon/The Fertilizer Institute/TFI - Programs/Research/4R Advocate Case Studies/4R Calculator/")
```

```{r packages}
library(knitr)
library(here)
library(readxl)
library(tidyverse)
```

```{r Load Advocate Inputs}
advocate1 = read_excel("4R Advocate Data/2025 Urich Farms/Inputs_MO.xlsx", sheet = 2)
```


```{r load calculator sheets}
#Machinery 
tractors = read_excel("Leanna's Calculator/Machinery.xlsx", sheet = 1)

machines = read_excel("Leanna's Calculator/Machinery.xlsx", sheet = 2)

implements = read_excel("Leanna's Calculator/Machinery.xlsx", sheet = 3)

equipment = read_excel("Leanna's Calculator/Machinery.xlsx", sheet = 5)

irrigation = read_excel("Leanna's Calculator/Machinery.xlsx", sheet = 4)

#Prices for Diesel
prices = read_excel("Leanna's Calculator/FertilizerPrices_Dec2025.xlsx", sheet = 2)
prices <- prices %>% #remove the duplicate rows that are being added at import
  distinct()
```

# Custom Costs 

```{r Custom Costs}
# Add a column for all custom costs per application 
df1 <- advocate1 %>% 
  mutate(totalCustomCost = appCostPerAc) 

```

# Drop purchase price columns if using standard price for machinery 
```{r}
df2 <- df1 %>%
  # Check if 'machPriceType' is "standard" and drop the specified columns -     other option is "own" meaning they provide their own cost 
  {if ("standard" %in% .$machPriceType) 
     select(., -machPurchPrice, -impPurchPrice) 
   else .}

```

# Check data types (DO NOT drop all empty columns - used in logic)
```{r}
str(df2)
```


# Machinery Self Application Costs 
Tact added to end of values associate with Tractors 
Imp added to end of columns associated with Implements 
Mach added to end of columns associated with Machines 

```{r}
# Join implement cost values 
df3 <- df2 %>%
  left_join(implements, by = c("implement", "impWdthFt"))
  # select(where(~ !all(is.na(.)))) - DO NOT drop columns. Keep same column template until end 

# Join tractor costs values
df4 <- df3 %>%
  left_join(tractors, by = c("machine", "tractorHp"),
            suffix = c("Imp", "Trac"))  #Imp (for implement) added to original columns
 
# Join machine cost values
df5 <- df4 %>%
  left_join(machines, by = c("machine"), # Join tractor costs by its HP class
            suffix = c("", "Mach")) #Imp (for implement) added to original columns

# Might result in empty columns, DO NOT drop columns. Keep same column template until end  
```

# Irrigation Costs 
```{r}
# Join irrigation cost values 
df6 <- df5 %>%
  left_join(irrigation, by = c("irrigation")) 

# Do not drop extra columns for calculation of cost as placeholder columns 
```

## Variable Costs 
### Labor Costs 
= labor($/hr) * Per Rate of implement (hr/ac)
1.	Per Rate (hr/ac) = 1/[speed * efficiency * (implement length/8.25)]

ii.	Center Pivot 
= labor($/hr) * labor (hrs/ac-in) * irrigation (ac-in) 

iii.	Drip irrigation 
= labor($/hr) * labor (hr/application) 

```{r Calculate Labor Costs}
# Cost for labor in $/hr 
Labor = 15 

df7 <- df6 %>%
  mutate(
    laborCost = case_when(
    is.na(irrigation) ~ perRateHrAc * Labor, # If not an irrigation app
    irrigation == "centerPivot" ~ Labor * laborHrAcIn * irrVolAcIn, 
    irrigation == "dripIrrigation"  ~  Labor * laborHrsApp, 
    TRUE ~ 0  # Default case if none of the above conditions are met
  ))
```

### Fuel Costs 
i. Tractor + Implement or Self-propelled machine 
= fuel use(gal/hr) * fuel cost * Per Rate of implement (hr/ac) 

ii. Center Pivot  
= fuel($/gal) * fuel (gal/ac-in) * irrigation (ac-in)  

iii. Drip irrigation  
None 

```{r Fuel}
# fuel cost global variable for irrigation (not sure where it comes from - could use average?)**
fuel_cost2 = 2.75

# Join prices df to get diesel costs for relevant years (3 yr avg)
df8 <- df7 %>%
  left_join(prices %>% select(Year, Diesel), by = c("year" = "Year"))

# Calculate Fuel Costs 
df9 <- df8 %>%
  mutate(fuelCost = case_when(
    is.na(irrigation) ~ fuelUseGalHr * perRateHrAc * Diesel,  # If irrigation is NA
    irrigation == "centerPivot" ~ fuel_cost2 * fuelGalAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    TRUE ~ 0  # Default case if none of the above conditions are met
  ))


```

### R&M Cost  
i. Tractor + Implement or Self-propelled machine 
= (tractor purchase price * R&M / useful life /annual hours) * Per Rate implement (hr/ac) +  
(Implement purchase price * R&M / useful life / annual hours) * Per Rate implement (hr/ac) 

ii. Center Pivot  
= R&M($/ac-in) * irrigation (ac-in)
```{r R&M}
df10 <- df9 %>%
  mutate(rmCost = case_when(
    is.na(irrigation) ~ ((machPurchPrice * rmTrac/ usefulLifeYrsTrac / annualUseHrsTrac) * perRateHrAc + 
                           (impPurchPrice * rmImp /usefulLifeYrsImp / annualUseHrsImp) * perRateHrAc),
    irrigation == "centerPivot" ~ rmDolAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    TRUE ~ 0)  # Default case if none of the above conditions are met
  )
```

## Ownership Costs 

### Depreciation 
  
i. Machine/Implement  
= (((tractor price – (tractor price * salvage)) / useful life / annual hours) * per rate of implement) +  
((implement price – (implement price * salvage)) / useful life / annual hours * per rate of implement))  

ii. Center Pivot  
= Ownership cost * irrigation (ac-in)  

iii. Drip irrigation
= ownership cost * length (ft) 

```{r Depreciation}
df11 <- df10 %>%
  mutate(deprecCosts = case_when(
    is.na(irrigation) ~ ((((machPurchPrice - (machPurchPrice * salvageTrac))/
                           usefulLifeYrsTrac / annualUseHrsTrac) * perRateHrAc) +
                        (((impPurchPrice - (impPurchPrice * salvageImp)) / 
                           usefulLifeYrsImp / annualUseHrsImp) * perRateHrAc)),  # If irrigation is NA
    irrigation == "centerPivot" ~ 
      ownershipDolAcIn * irrVolAcIn,  # If irrigation is "centerPivot"
    irrigation == "dripIrrigation" & !is.na(irrigation) ~ 
      ((ownershipDolFt * impWdthFt) * (irrVolAcIn/totalWater)),
    TRUE ~ 0)) 
```
### Interest 

i. Machine/Implement 
= (((machinery price+(machinery price * salvage))/2* interest rate / annual hours * per rate of implement) + 
((implement price +(implement price * salvage))/2* interest rate / annual hours * per rate of implement))

ii. Center pivot  
None 

iii. Drip irrigation  
None

```{r Interest}
interest_rate = 0.05

df12 <- df11 %>%
  mutate(interestCosts = case_when(
    is.na(irrigation) ~ 
      (((machPurchPrice + (machPurchPrice * salvageTrac)) / 2 * 
          interest_rate / annualUseHrsTrac * perRateHrAc) +
         ((impPurchPrice+(impPurchPrice*salvageImp))/2*interest_rate/annualUseHrsImp *
                            perRateHrAc)),
    TRUE ~ 0) # Default case if none of the above conditions are met
) 

```

## Total Machinery Costs 
```{r Total Machinery Costs}
df13 <- df12 %>% 
  # mutate(# Add dummy columns for values in equations when farmer doesn't use their own machinery  
  #   if (!"TotalCustomCost" %in% colnames(df)) {df <- df %>% mutate(TotalCustomCost = NA)}) %>%
  mutate(variableCosts = 
           coalesce(laborCost, 0) + 
           coalesce(fuelCost,0) + 
           coalesce(rmCost, 0))%>%
  mutate(ownershipCosts = 
           coalesce(interestCosts, 0) + 
           coalesce(deprecCosts, 0)) %>%
  mutate(machineryCosts = 
           coalesce(variableCosts, 0) + 
           coalesce(ownershipCosts, 0) + 
           coalesce(totalCustomCost, 0))

  
inputs1 <- df13 %>%
  select(-c(31:70)) 

# drop columns for calculating costs. 31 columns in original df (advocate 1) and need the last 3 columns 
  
```


# Write File 
```{r Write File}
# Save csv
write.csv(inputs1, file = "4R Advocate Data/2025 Urich Farms/Inputs_machCosts.csv", row.names = FALSE)
```

```{r}
rm(list = ls()) # clear environment
```
