---
title: "source_costs"
output: html_document
date: "2024-12-02"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set wd}
setwd("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/4R Advocate Case Studies/")
```

```{r load advocate data}
library(readxl)

advocate = read.csv("2024 Bryant Lowe/Outputs/DE Inputs_machCosts.csv")
advocate = read.csv("2024 Mason Roberts/Outputs/GA Inputs_machCosts.csv")
```


```{r load calculator sheets}
#Sources 
prices = read_excel("C:/Users/LeannaLeverich/The Fertilizer Institute/TFI - Programs/Research/4R/4R Calculator/Leanna's Calculator/FertilizerPrices_Dec2024.xlsx", sheet = 2)
prices <- prices %>% #remove the duplicate rows that are being added at import
  distinct()
```

```{r}
 # Load the tidyr package
library(tidyr)

# Convert the dataframe to long format
pricesLong <- prices %>%
  select(where(~ !all(is.na(.)))) %>% 
  pivot_longer(
    cols = starts_with("AA"):starts_with("K"),  # Select the columns to pivot
    names_to = "source",  # New column for the names of the subjects
    values_to = "sourceCost")    # New column for the scores 

# View the long-format dataframe
print(pricesLong)
```



# Specific product prices 

ii.	Specific Sources
  1.	Tons
    = (Price of source/2000) * source rate 
  2.	lbs
    = Price of source * source rate

```{r}
# Join prices to inputs for specific sources prices 
df1 <- advocate %>%
  left_join(pricesLong, by = c("source", "year"= "Year")) %>%
  mutate(sourceCost = (sourceCost/2000) * sourceRate)

df2 <- df1 %>%
  left_join(prices, by = c("year" = "Year"), # Join all columns by year for prices 
            suffix = c("perc", "price"))  # add a suffix to the columns for the 2 dataframes being joined 
  
# %>% select(where(~ !all(is.na(.)))) #Drop all columns where all rows are NA
```

# Source Costs 
i.	Custom Blends or Starter 
    1.	lbs  
      = (Price of N that year * N percentage * source rate) + (Price of P that year * P percentage * source rate) + (price         of K that year * K percentage * source rate)
    2.	Tons
      = (Price of N that year * N percentage * source rate * 2000) + (Price of P that year * P percentage * source rate *         2000) + (price of K that year * K percentage * source rate * 2000)

```{r}
# Calculate prices for blends/starter 
df3 <- df2 %>%
  mutate(sourceCost = case_when(
    source == "blend" | source == "starter"  ~ 
      (Nprice * (Nperc/100) * sourceRate) + (Pprice * (Pperc/100) * sourceRate) + (Kprice * (Kperc/100) * sourceRate), 
    TRUE ~ sourceCost  # Default case: retain the current value of "price"
  )) 

```

Calculate source costs for inhibitors, micros, other

```{r}
df4 <- df3 %>%
  mutate(microCost = (inhibitorDolTon/2000) * sourceRate, 
         inhibitorCost = (inhibitorDolTon/2000) * sourceRate, 
         otherFertCost = (otherFertDolTon/2000)* sourceRate) %>%
  mutate(specialSourceCost = 
           (coalesce(microCost, 0) + 
           coalesce(inhibitorCost, 0) + 
           coalesce(otherFertCost, 0))) 

df4$specialSourceCost

df5 <- df4 %>%
  mutate(totalSourceCost = sourceCost + specialSourceCost, 
         totalCostPerApp = machineryCosts + totalSourceCost)
         
```


Clean up dataframe
```{r}
df6 <- df5 %>%
  select(-c(33:49))%>%
  # select(-starts_with("AA") & !starts_with("valueAdd"))%>% # redundant code 
  select(where(~ !all(is.na(.))))
```

Check total $/ac for each field and year
```{r}
# Summarize costs for total costs for machinery and source by year and field
costSummary <- df6 %>%
  group_by(year, field) %>%
  summarize(
    totalMachineCost = sum(machineryCosts), 
    totalSourceCost = sum(sourceCost), 
    totalCost = sum(totalCostPerApp)
)

```

```{r}
# Save csv
write.csv(df6, file = "2024 Bryant Lowe/Outputs/DE Inputs_sourceMachCosts.csv", row.names = FALSE)
write.csv(df6, file = "2024 Mason Roberts/Outputs/GA Inputs_sourceMachCosts.csv", row.names = FALSE)

```

